---
title: "Lab 5"
author: "keenan schissel"
date: "2/23/2021"
output: html_document
---

Packages
```{r}
library(tidyverse)
```

Rename each variable so that it is all lower-case and uses an underscore _ instead of a period . in the name (the recommended coding style in the tidyverse style guide). Print the resulting table.
```{r}
Q1<-rename(iris,sepal_length=Sepal.Length,sepal_width=Sepal.Width,petal_length=Petal.Length,petal_width=Petal.Width,species=Species)
Q1
```

Convert the four numerical variables from cm to mm by multiplying by 10. Print the resulting table.
```{r}
Q2<-select(Q1, sepal_length, sepal_width, petal_length, petal_width)
Q2
mutate(Q1, Q2 = Q2 * 10)
```

Calculate sepal area and petal area (area is equal to length multiplied by width). Print a table with only the variables sepal area, petal area, and species.
```{r}
Q3 <-mutate(Q1, sepal_area=sepal_length*sepal_width,petal_area=petal_length*petal_width)
Q3
Q3.5 <-select(Q3,sepal_area,petal_area)
Q3.5
```

Calculate the following statistics for the entire dataset from the sepal length variable and print the resulting table:
```{r}
summarize(
  Q2,
  sampl_size = n(),
  max=max(sepal_length),
  min=min(sepal_length),
  range=range(sepal_length),
  median=median(sepal_length),
  q1=quantile(sepal_length, probs = 0.25),
  q2=quantile(sepal_length, probs = 0.75),
  IQR=IQR(sepal_length)
)
```

Calculate the following statistics for each species from the petal width variable and print the resulting table:
```{r}
Q5<-group_by(Q1,species)

iris_sum<-summarize(
  Q5,
  mean=mean(petal_width),
  sampl_size = n(),
  sd=sd(petal_width),
  var=var(petal_width),
  sem=sd / sqrt(sampl_size), 
  ci_upper=mean+2 * sem,
  ci_lower=mean-2 * sem)
iris_sum
```

Visualize the relationship between petal length and species using a strip plot.
```{r}
ggplot(data = Q1) + 
  geom_jitter(mapping = aes(x=species,y=petal_width))
```

Starting with the previous graph, add the mean and 95% confidence interval for each species
```{r}
ggplot(data=Q1) +
  geom_jitter(mapping=aes(x=species,y=petal_width)) +
  geom_crossbar(
    data=iris_sum,
    mapping=aes(
      x=species,
      y=mean,
      ymax=ci_upper,
      ymin=ci_lower),
    color="red"
  )
```

Visualize the relationship between petal length, petal width, and species using a scatterplot. Map the two numerical variables to the x and y axes and map species to the color and shape aesthetics.
```{r}
ggplot(data=Q1) +
  geom_point(mapping=aes(x=petal_length,y=petal_width,color=species))
```

```{r session-info}
sessioninfo::session_info()
```